Parameter estimation

### 2022-05-31 SFE-report/Paramter_fitting.tex
31.05

\begin{equation}
	P\{A_i|B\} = \cfrac{P\{B|A_i\}P\{A_i\}}{P\{B\}}
\end{equation}

@@

\begin{equation} \label{EQ: Bayes_discrit}
	P\{A_i|B\} = \cfrac{P\{B|A_i\}P\{A_i\}}{ \sum_{i=1}^{n} P\{B|A_i\} P\{A_i\} }
\end{equation}

@@

\begin{equation} \label{EQ: Bayes_continous}
	p\left( \theta|X=\textbf{x} \right) = p\left( \theta|\textbf{x} \right) = \cfrac{L\left( \theta|\textbf{x} \right) p(\theta)}{\int_{-\infty}^{+\infty} L\left( \theta|\textbf{x} \right) p(\theta) d\theta}
\end{equation}

@@

\begin{equation}
	\int_{-\infty}^{+\infty} p\left( \textbf{x}|\theta \right) p(\theta) d\theta = p(\textbf{x})
\end{equation}

@@

\begin{equation}
	p\left( \theta|\textbf{x} \right) = \cfrac{L\left( \theta|\textbf{x} \right)}{\int_{-\infty}^{+\infty} L\left( \theta|\textbf{x}\right) d\theta}
\end{equation}

@@

	\begin{equation}
		P\left( A_1, \text{or~} A_2,..., \text{or~} A_n \right) = \sum_{i=1}^{n} P\left( A_i \right)
	\end{equation}

	Very often we let
	\begin{equation}
		\sum_{i=1}^{n} P\left( A_i \right) = 1
	\end{equation}

@@

	\begin{equation}
		\sum_{i=1}^{n} q = nq = 1 \qquad \text{or} \qquad  q=\cfrac{1}{n} = P\left( A_i \right)
	\end{equation}

@@

	\begin{equation}
		P\left( A_1 \cup A_2 \cup ... \cup A_n \right) = P(A_1) + P(A_2) + ... + P(A_n)
	\end{equation}

@@

	\begin{equation}
		P\left( A \text{~and~} B\right) = P(A)P(B)
	\end{equation}

@@

	\begin{equation} \label{EQ: Probabiltiy_independet_a}
		P\left( A \cap B  \right) = P(A)P(B)
	\end{equation}

@@

	\begin{equation} \label{EQ: Probabiltiy_independet_b}
		P\left(A|B\right) = \cfrac{P\left( A \cap B \right)}{P(B)}
	\end{equation}

@@

	 \begin{subequations}
		\begin{alignat}{2}
			P\left( A \cap B \right) &= P(B) P(A|B) \label{EQ: Probabiltiy_dependet_a} \\
			&= P(A)P(B|A) \label{EQ: Probabiltiy_dependet_b}
		\end{alignat}
	\end{subequations}

@@

	\begin{equation}
		\begin{split}
			P(A_1 \text{~and~} A_2 \text{~and~} ... \text{~and~} A_n) &= P(A_1)\cdot P(A_2) \cdot  ...\cdot P(A_n) \\ &= \prod_{i=1}^{n} P(A_i)
		\end{split}
	\end{equation}

@@

	\begin{equation}
		P(A) + P(B) - P(A \cap B) = P(A \cup B)
	\end{equation}

@@

\begin{equation}
	Y(t) = y(t) + \epsilon(t)
\end{equation}

@@

\begin{equation}
	\textbf{Y}(t_i) = h(t_i)y(t_i) + \epsilon(t_i) \qquad 1 \leq i \leq n
\end{equation}

where $\textbf{Y}(t_i)$ is an $n \times 1$ column vector, $h(t_i)$ is an $n \times v$ matrix given \textit{a priori}, and $\epsilon (t_i)$ is an $x \times 1$ column vector ("noise" vector) whose elements are the unobservable errors. The above equation can be written in general form as:

\begin{equation} \label{EQ: Measurment_noise}
	\textbf{Y}(t_i) = \Psi(\theta, y_0, t_i) + \epsilon(t_i)
\end{equation}
where $\Psi$ is the matrix of functions.

@@

\begin{equation}
	\cfrac{d\textbf{Y}}{dt} = \cfrac{dy}{dt} + \epsilon
\end{equation}

@@

\begin{equation}
	\phi =  \cfrac{1}{2} \int_{0}^{t_f}  \left[ \textbf{Y} - \Psi \right]^T \Gamma^{-1} \left[ \textbf{Y} - \Psi \right] dt
\end{equation}

where $\Gamma$ is the covariance matrix (or perhaps a matrix of appropriate weights), and $\phi$ is the time integrated value of the error squared ("integral squared error"). If the observations are made at discrete instants of time, $t_i$, $i = 1,2,...,n$, the Markov criterion is to minimize:

\begin{equation}
	\phi = \cfrac{1}{2} \sum_{i=1}^{n} \left[ \textbf{Y}(t_i) - \Psi(t_i) \right]^T \Gamma^{-1} \left[ \textbf{Y}(t_i) - \Psi(t_i) \right] 
\end{equation}

@@

@@

\begin{equation} \label{EQ: Posterior_distribution}
	p\left( \theta,y_0|y \right) = \cfrac{p\left( \theta, y_0, y \right)}{p\left( y \right)}
\end{equation}

@@

\begin{equation}
	\begin{split}
		p\left( \theta,y_0,y \right) &= p\left( y(t_n)|\theta,y_0,y(t_{n-1}),...,y(t_1) \right) \\ &\cdot p\left( \theta,y_0,y(t_{n-1}),...,y(t_1) \right)
	\end{split}
\end{equation}

@@

\begin{equation}
	p\left( \theta,y_0,y \right) = p\left( \theta,y_0 \right) \prod_{i=1}^{n} p\left( \theta,y_0|y \right)
\end{equation}

@@

\begin{equation}
	p\left( y(t_i)|\theta,y_0, y(t_{i-1}),...,y(t_1) \right) = p\left( y(t_i)|\theta,y_0 \right)
\end{equation}

@@

\begin{equation}
	p\left( \theta,y_0|y \right) = \cfrac{p\left( \theta,y_0 \right) \prod_{i=1}^{n} p\left( y(t_i)|\theta,y_0 \right) }{p\left( y \right)}
\end{equation}

@@

\begin{equation} \label{EQ: MLE_Lagrangian}
	\begin{split}
		L^* \equiv \ln p\left( \theta,y_0 \right) + \sum_{i=1}^{n} \left\{ \ln p(y(t_i)|\theta,y_0) \right. \\
		\left. + \lambda^T \cdot \left[ y(t_i) - \Psi\left( \theta,y_0,t_i \right) - \epsilon(t_i) \right] \right\} \\ - \ln \left[ p\left( y \right) \right]
	\end{split}
\end{equation}

>>>>

{\footnotesize
\begin{equation}
	P\{A_i|B\} = \cfrac{P\{B|A_i\}P\{A_i\}}{P\{B\}}
\end{equation} }

@@

{\footnotesize
\begin{equation} \label{EQ: Bayes_discrit}
	P\{A_i|B\} = \cfrac{P\{B|A_i\}P\{A_i\}}{ \sum_{i=1}^{n} P\{B|A_i\} P\{A_i\} }
\end{equation} }

@@

{\footnotesize
\begin{equation} \label{EQ: Bayes_continous}
	p\left( \theta|X=\textbf{x} \right) = p\left( \theta|\textbf{x} \right) = \cfrac{L\left( \theta|\textbf{x} \right) p(\theta)}{\int_{-\infty}^{+\infty} L\left( \theta|\textbf{x} \right) p(\theta) d\theta}
\end{equation} }

@@

{\footnotesize
\begin{equation}
	\int_{-\infty}^{+\infty} p\left( \textbf{x}|\theta \right) p(\theta) d\theta = p(\textbf{x})
\end{equation} }

@@

{\footnotesize
\begin{equation}
	p\left( \theta|\textbf{x} \right) = \cfrac{L\left( \theta|\textbf{x} \right)}{\int_{-\infty}^{+\infty} L\left( \theta|\textbf{x}\right) d\theta}
\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		P\left( A_1, \text{or~} A_2,..., \text{or~} A_n \right) = \sum_{i=1}^{n} P\left( A_i \right)
	\end{equation} }

	Very often we let

	{\footnotesize
	\begin{equation}
		\sum_{i=1}^{n} P\left( A_i \right) = 1
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		\sum_{i=1}^{n} q = nq = 1 \qquad \text{or} \qquad  q=\cfrac{1}{n} = P\left( A_i \right)
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		P\left( A_1 \cup A_2 \cup ... \cup A_n \right) = P(A_1) + P(A_2) + ... + P(A_n)
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		P\left( A \text{~and~} B\right) = P(A)P(B)
	\end{equation} }

@@

	{\footnotesize
	\begin{equation} \label{EQ: Probabiltiy_independet_a}
		P\left( A \cap B  \right) = P(A)P(B)
	\end{equation} }

@@

	{\footnotesize
	\begin{equation} \label{EQ: Probabiltiy_independet_b}
		P\left(A|B\right) = \cfrac{P\left( A \cap B \right)}{P(B)}
	\end{equation} }

@@

	{\footnotesize
	 \begin{subequations}
		\begin{alignat}{2}
			P\left( A \cap B \right) &= P(B) P(A|B) \label{EQ: Probabiltiy_dependet_a} \\
			&= P(A)P(B|A) \label{EQ: Probabiltiy_dependet_b}
		\end{alignat}
	\end{subequations} }

@@

	{\footnotesize
	\begin{equation}
		\begin{split}
			P(A_1 \text{~and~} A_2 \text{~and~} ... \text{~and~} A_n) &= P(A_1)\cdot P(A_2) \cdot  ...\cdot P(A_n) \\ &= \prod_{i=1}^{n} P(A_i)
		\end{split}
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		P(A) + P(B) - P(A \cap B) = P(A \cup B)
	\end{equation} }

@@

{\footnotesize
\begin{equation}
	Y(t) = y(t) + \epsilon(t)
\end{equation} }

@@

{\footnotesize
\begin{equation}
	\textbf{Y}(t_i) = h(t_i)y(t_i) + \epsilon(t_i) \qquad 1 \leq i \leq n
\end{equation} }
where $\textbf{Y}(t_i)$ is an $n \times 1$ column vector, $h(t_i)$ is an $n \times v$ matrix given \textit{a priori}, and $\epsilon (t_i)$ is an $x \times 1$ column vector ("noise" vector) whose elements are the unobservable errors. The above equation can be written in general form as:

{\footnotesize
\begin{equation} \label{EQ: Measurment_noise}
	\textbf{Y}(t_i) = \Psi(\theta, y_0, t_i) + \epsilon(t_i)
\end{equation}}
where $\Psi$ is the matrix of functions.

@@

{\footnotesize
\begin{equation}
	\cfrac{d\textbf{Y}}{dt} = \cfrac{dy}{dt} + \epsilon
\end{equation} }

@@

{\footnotesize
\begin{equation}
	\phi =  \cfrac{1}{2} \int_{0}^{t_f}  \left[ \textbf{Y} - \Psi \right]^T \Gamma^{-1} \left[ \textbf{Y} - \Psi \right] dt
\end{equation} }
where $\Gamma$ is the covariance matrix (or perhaps a matrix of appropriate weights), and $\phi$ is the time integrated value of the error squared ("integral squared error"). If the observations are made at discrete instants of time, $t_i$, $i = 1,2,...,n$, the Markov criterion is to minimize:

{\footnotesize
\begin{equation}
	\phi = \cfrac{1}{2} \sum_{i=1}^{n} \left[ \textbf{Y}(t_i) - \Psi(t_i) \right]^T \Gamma^{-1} \left[ \textbf{Y}(t_i) - \Psi(t_i) \right] 
\end{equation} }

@@

\subsection{Cumulative data}

One characteristic feature of certain special experiments is the use of the same batch of materials for the entire series of measurements. If unobservable error in the first observation is designated as $\epsilon_1$, the second error observation $\epsilon_2'$ includes $\epsilon_1$ plus a random component introduced side from $\epsilon_1$, or $\epsilon_2' = \epsilon_1 + \epsilon_2$. The error in the third observation is $\epsilon_3' = \epsilon_1 + \epsilon_2 + \epsilon_3$, and so forth. \citet{Mandel1957} distinguished between the usually assumed type of independent measuring error in the dependent variable and a 'cumulative' or interval error in which each new observation includes the error of the previous observations. Cumulative errors, arising because of the fluctuations as a function of time in the process itself due to small changes in operating, are not independent - only the differences in measurment from one period to the next are independent. Thus, if we consider two model:

{\footnotesize
\begin{flalign*}
	&\text{Model A - independent error:} &&\nonumber \\
	Y_i &= \Psi(\theta, y_0, t_i) + \epsilon_i \qquad i = 1,2,...,n&& \\
	&\text{Model B - cumulative data:}&& \nonumber \\
	Y_j - Y_{j-1} &= \Psi(\theta, y_0, t_j) - \Psi(\theta, y_0, t_{j-1}) + \epsilon_j, \quad j = 1,2,...,n&& \\
	&\text{or}&& \nonumber \\
	Y_i = \sum_{j=1}^{i} &\left( Y_j - Y_{j-1} \right) = \sum_{j=1}^{i} \left( \Psi(\theta, y_0, t_j) - \Psi(\theta, y_0, t_{j-1}) \right) + \sum_{j=1}^{i} \epsilon_j &&
\end{flalign*} }

@@

{\footnotesize
\begin{equation} \label{EQ: Posterior_distribution}
	p\left( \theta,y_0|y \right) = \cfrac{p\left( \theta, y_0, y \right)}{p\left( y \right)}
\end{equation} }

@@

{\footnotesize
\begin{equation}
	\begin{split}
		p\left( \theta,y_0,y \right) &= p\left( y(t_n)|\theta,y_0,y(t_{n-1}),...,y(t_1) \right) \\ &\cdot p\left( \theta,y_0,y(t_{n-1}),...,y(t_1) \right)
	\end{split}
\end{equation} }

@@

{\footnotesize
\begin{equation}
	p\left( \theta,y_0,y \right) = p\left( \theta,y_0 \right) \prod_{i=1}^{n} p\left( \theta,y_0|y \right)
\end{equation} }

@@

{\footnotesize
\begin{equation}
	p\left( y(t_i)|\theta,y_0, y(t_{i-1}),...,y(t_1) \right) = p\left( y(t_i)|\theta,y_0 \right)
\end{equation} }

@@

{\footnotesize
\begin{equation}
	p\left( \theta,y_0|y \right) = \cfrac{p\left( \theta,y_0 \right) \prod_{i=1}^{n} p\left( y(t_i)|\theta,y_0 \right) }{p\left( y \right)}
\end{equation} }

@@

{\footnotesize
\begin{equation} \label{EQ: MLE_Lagrangian}
	\begin{split}
		L^* \equiv \ln p\left( \theta,y_0 \right) + \sum_{i=1}^{n} \left\{ \ln p(y(t_i)|\theta,y_0) \right. \\
		\left. + \lambda^T \cdot \left[ y(t_i) - \Psi\left( \theta,y_0,t_i \right) - \epsilon(t_i) \right] \right\} \\ - \ln \left[ p\left( y \right) \right]
	\end{split}
\end{equation}}
